An Abstraction-based Approach to Network Synthesis

### 2016-11-03 paper/abstraction/paper.tex
stashing changes

% These networks often
% host a myriad of services and every second of downtime imposes significant
% financial cost or disruption to consumers.  
Such networks typically have uniform, tree-shaped topologies that involve
hundreds or even thousands of devices.  Since it is not possible for humans
to think about configuring each individual device independently, 
network operators typically classify the devices, assigning them a particular
\emph{role} such as ``top-of-rack switch'' or ``tier-one aggregation switch.''
Each such role is codified as a \emph{configuration template}---essentially
a macro that may be instantiated with different concrete values to
generate a configuration for each device that plays the role.

Such templates have no particular semantics other than the semantics
that one obtains from instantiating a template's parameters.  Moreover,
template construction is a manual process:  network engineers do their
best to design templates by hand that will work effectively in the context 
of their particular network topology.  However, there is no support for
validating key properties such as connectivity and fault tolerance, 
aside from instantiating the templates and testing.  In addition,
network engineers have no tools to help them design for network
evolution.  For instance, the process of adding new equipment to
expand the capacity of data center may not be compatible with the original
templates, or may cause a trickle-down of changes to be made network-wide.
Unfortunately, each change to a router configuration typically incurs
some risk of temporary instable in a network, so minimizing such changes
is an important design goal.  It is no wonder that a major
network outage, such as the one that grounded all United Air flights
in July 2015~\cite{x}, makes international news every few months.

To combat these problems, in this paper, we present \sysname, a 
new high-level language designed to 
provide improved support for the configuration of large, structured networks.
\sysname allows network operators to define abstract network topologies that
classify devices in terms of abstract roles and specify the connectivity
invariants satisfied by each role.  For instance, a programmer may
define separate roles for ``top-of-rack switch'' and 
``tier-one aggregation switch'' and specify that every top-of-rack switch
is connected to at least two different tier-one aggregation switches.

In order to specify \emph{routing policy} for such abstract networks,
\sysname borrows notation from the recently-defined \propane
programming language~\cite{propane}.  However, whereas \propane policies
refer to individual devices, \sysname policies allows users to
refer to abstract roles.  The \sysname compiler compiles these role-based
policies into templates, with one template per role, and these templates
may be instantiated to generate concrete configurations.

@@

To summarize, the central contribution of this paper are:

\begin{enumerate}
\item Design of new network topology abstraction mechanisms involving roles and
connectivity invariants.
\item Design of two new algorithms for analyzing abstract networks: one which
guarantees correct compilation and one which determines fault tolerance 
properties of network policy.  Both algorithms provide guarantees for any 
concrete network satisfying the abstract connectivity invariants.
\item Design of template generation algorithms that support network
evolution with minimal change to existing devices.
\item Implementation and evaluation of a \sysname compiler that generates 
configuration templates and is up to two orders of magnitude faster 
than its closest competitor on large networks.
\end{enumerate}

@@

\section{Motivation}
\label{sec:motivation}

@@

%We motivate \sysname using an example data center network. Our network, 
Consider the data center network shown in Figure~\ref{fig:example}.  The boxes in the diagram
represent \emph{switches}, and these switches play several different roles:  T[1--8] denote top-of-rack (ToR) switches, A[1--8] denote aggregation switches, and S[1--2] denote spine switches. The spine switches are connected to the
rest of the internet; any traffic to or from the outside world flows through them.  The aggregation switches
serve as a bridge between spine and ToR switches.   Finally, each ToR switch is
connected to a set servers (``a rack''): P[1--8] denote groups of addresses for services hosted in each rack. Such groups are called address {\em prefixes} because they are encoded using the adresses' shared prefix.

We assume that each switch is running an instance of the \emph{Border Gateway Protocol} (BGP), 
as is common in large data centers. In this protocol, adjacent switches use BGP to exchange messages
with one another concerning the availability of routes to destinations.  A BGP \emph{configuration}
(one for each switch) dictates which offered routes a switch will select and how to propagate chosen routes to its 
neighbours.\footnote{If this submission is accepted and more space is available in the final version, 
we will give a detailed explanation of BGP for the PLDI audience.  Henceforth, we assume knowledge of basic BGP
mechanics.}

@@

\subsection{Limitations of concrete synthesis}

@@

\subsection{Limitations of today's template systems}

@@

\subsection{Product graph distributed configurations}

>>>>

% These networks often
% host a myriad of services and every second of downtime imposes significant
% financial cost or disruption to consumers.  
Such networks typically have uniform, tree-shaped topologies that involve
hundreds or even thousands of devices.  Since it is not possible for humans
to think about configuring each individual device independently, 
network operators typically classify the devices, assigning them a particular
\emph{role} such as ``top-of-rack switch'' or ``tier-one aggregation switch.''
Each such role is codified as a \emph{configuration template}---a macro 
that may be instantiated with different concrete values to
generate a configuration for each device that plays the role.

Unfortunately, such templates have no particular semantics other than the 
semantics
that one obtains from instantiating a template's parameters.  
\dpw{add blah blah about low-level, assembly language, device by device}
Moreover,
template construction is a manual process:  network engineers do their
best to design templates that will work effectively in the context 
of their particular network topology.  However, there is no support for
validating key properties of an assembly of templates, such as connectivity 
and fault tolerance, 
aside from instantiating the templates and testing.  In addition,
network engineers have no tools to help them design for network
evolution.  For instance, the process of adding new equipment to
expand the capacity of data center may not be compatible with the original
templates, or may cause a trickle-down of changes to be made network-wide.
Each change to a router configuration typically incurs
some risk of temporary instable in a network, so minimizing such changes
is an important design goal.  It is no wonder that a major
network outage, such as the one that grounded all United Air flights
in July 2015~\cite{x}, makes international news every few months.

To combat these problems, we present \sysname, a 
new high-level language designed to 
provide improved support for the configuration of large, structured networks.
\sysname allows network operators to define \emph{abstract network topologies} 
that
classify devices in terms of their roles and how they are connected
to one another.  For instance, a programmer may
define separate roles for ``top-of-rack switch'' and 
``tier-one aggregation switch'' and specify that every top-of-rack switch
is connected to at least two different tier-one aggregation switches,
thereby providing a degree of fault tolerance of one aggregation switch
were to fail.

To specify \emph{routing policy} for such abstract networks,
\sysname borrows notation from the recently-defined \propane
programming language~\cite{propane}.  As such, it is able to
specify the paths that traffic should flow along at a high-level of
abstraction.  However, whereas \propane policies
refer to paths of individual devices, \sysname policies 
refer to paths of abstract roles.  The \sysname compiler compiles such
policies into templates, with one template per role, and these templates
may be instantiated to generate concrete configurations.

@@

To summarize, the central contributions of this paper are:

\begin{enumerate}
\item Design of new topology abstractions involving roles and
connectivity invariants for network programmers.
\item Design of two new algorithms for analyzing routing policy over 
abstract networks: one which
guarantees correct compilation and one which determines fault tolerance 
properties of network policy.  Both algorithms provide guarantees for any 
concrete network satisfying the abstract connectivity invariants.
\item Design of template generation algorithms that support network
evolution.
\item Implementation and evaluation of a \sysname compiler that generates 
configuration templates and is up to two orders of magnitude faster 
than its closest competitor on large networks.
\end{enumerate}

@@

\section{Background and Motivation}
\label{sec:motivation}

@@

\paragraph*{Background}  Traditional routers use 
distributed protocols to exchange messages about available routes. 
One such protocol is the Border Gateway Protocol (BGP).  BGP is the
standard for inter-domain routing (the act of exchanging routing information 
between independently-owned networks) and is also commonly used in
data centers because it allows operators to define relatively flexible
routing policy while also scaling effectively to large networks.
When used within a data center, each router is configured with its own \emph{AS number},
which identifies it.

Each BGP message (also called a \emph{route announcement}) contains a 
\emph{destination prefix}, such as 10.1.1.0/24,\footnote{An address
such as 10.1.1.0 is 32 bits, with each of the four values representing 8 bits.  
A prefix such
as 10.1.1.0/24 denotes the set of addresses that begin with the first
24 bits of 10.1.1.0, \IE, all addresses that begin with 10.1.1.} as well as an
\emph{AS path}, such as XYZ (where each of X, Y and Z would be an AS number).  
Such messages indicate that the set of addresses
denoted by the prefix may be reached by following the path XYZ. 
A BGP message may also contain additional attributes, such as 
\emph{local preference} attribute and one or more
\emph{community values}.

A BGP router may receive messages about the same prefix from several
of its \emph{peers} (\IE, neighbouring routers).  When this occurs,
it processes the messages as follows:
\begin{enumerate}
\item It \emph{filters} (drops) some of the incoming messages based on
any their attributes (AS path, community values, peer it came from, \ETC),
and optionally modifies attributes of the messages it does not drop.
This is usually called the \emph{import} stage.
\item From the remaining messages, it selects the one message it
most \emph{prefers}. Selection is based on a number of criteria, but the
most important is the local preference number: Higher local preferences
are preferred.  Ties are broken by AS path length --- BGP defaults to 
shortest path routing in such cases.
\item Finally, it \emph{exports} the preferred message to selected neighbors,
adding its own AS number to the path and optionally adding or
modifying attributes.
\end{enumerate}
When a BGP router is configured, the user tunes the import and export
stages of the above process, choosing which messages to import, how to
modify them and how to export them.  By modifying the local preferences
on messages in the inport phase, the user determines which messages will
be selected.  This in turn determines how traffic destined for the given
address will be routed.  Keep in mind that if BGP messages to a destination
are propagated in one direction (left-to-right), then traffic will wind
up flowing in the opposite direction (right-to-left).

Rather than export 2$^n$ addresses in the same range, routers will
often perform \emph{route aggregation}, and export a single (32-$n$)-bit
prefix.  During aggregation, whenever an advertisement for 
\emph{any} of the 2$^n$ addresses
reaches the router, the \emph{entire} aggregate prefix will be announced.
While useful for decreasing announcements and avoiding churn, such
aggregation can be dangerous:  If a message for one of the 2$^n$ addresses (X)
reaches the aggregation router, but another one of the addresses (Y)
cannot reach the aggregation router due to a failure then the aggregate
will be announced, possibly drawing traffic to Y, even though Y cannot be
reached.  This is known as an \emph{aggregation-induced black hole}~\cite{route-aggregation}.
Networks should be configured to be sufficiently fault tolerant to minimize
the chances of such black holes.

\paragraph*{An Example}
%We motivate \sysname using an example data center network. Our network, 
Consider the data center network shown in Figure~\ref{fig:example}.  The boxes in the diagram
represent \emph{switches}, and these switches play several different roles:  T[1--8] denote top-of-rack (ToR) switches, A[1--8] denote aggregation switches, and S[1--2] denote spine switches. The spine switches are connected to the
rest of the internet; any traffic to or from the outside world flows through them.  The aggregation switches
serve as a bridge between spine and ToR switches.   Finally, each ToR switch is
connected to a set servers (``a rack''): P[1--8] denote groups of addresses for services hosted in each rack. 
% Such groups are called address {\em prefixes} because they are encoded using the adresses' shared prefix.

%% We assume that each switch is running an instance of the \emph{Border Gateway Protocol} (BGP), 
%% as is common in large data centers. In this protocol, adjacent switches use BGP to exchange messages
%% with one another concerning the availability of routes to destinations.  A BGP \emph{configuration}
%% (one for each switch) dictates which offered routes a switch will select and how to propagate chosen routes to its 
%% neighbours.\footnote{If this submission is accepted and more space is available in the final version, 
%% we will give a detailed explanation of BGP for the PLDI audience.  Henceforth, we assume knowledge of basic BGP
%% mechanics.}

@@

\paragraph*{Limitations of concrete synthesis}

@@

\paragraph*{Limitations of today's template systems}

@@

\subsection{Compilation to configurations}
